\section{Introduction}

Traditional approaches to human identification such as PIN numbers, tokens and
passwords used in magnetic credit cards are considered obsolete these days by
the ease in which they can be circumvented. Biometrics comes in as a solution to
overcome traditional identification techniques, since biological characteristics
can not be forged. Among the physical characteristics of a human being, the face
is the most used modality \cite{delac2004survey}. It has a good acceptance in
recognition systems and since the emergence of photography, government entities
and private organizations have been maintaining databases with photographs of
individuals for their personal identification.

Therefore, face recognition (FR) has gained a great deal of attention in the
academic and industrial communities. Many different methods used for solving
this problem have been proposed and tested
\cite{turk1991eigenfaces,sirovich1987low,belhumeur1997eigenfaces}. From these it
has become evident that one of the most important issue is feature extraction.
In the real word, face images are usually influenced by variances such as
illuminations, posture, occlusions, and expressions. Additionally, there is fact
that the difference from the same person would be much larger than that from
different people. Thus, finding numerical descriptors which are efficient and
discriminant enough becomes a key in performing facial recognition sufficiently
well.

Given a human face image, one could intuitively try to identify distinct
features such as eyes, mouth, nose, or chin. By measuring their relative
positions, distinguishing between different people may be possible. However,
rather than performing FR with the prior described \textit{Geometric-based}
approach, one could leverage Machine Learning in representing the images as 2D
arrays of grayscale intensities and trying to find underlying patterns in the
data. 

One of the first and well-known method used is the linear principal component
analysis (PCA), also known as Eigenfaces, which was originally proposed by
Sirovich and Kirby \cite{sirovich1987low}, and then developed by Turk and
Pentland \cite{turk1991eigenfaces}. Currently, deep learning including deep
neural network has shown its great success on image expression
\cite{krizhevsky2012imagenet}. Here, I will investigate and compare the new
approach to the FR problem in the form of an autoencoder. Autoencoders are a
nonlinear multilayer structure extracting the features of the image, step by
step, in an unsupervised way. The succeeding hidden layers are responsible for
the more and more compressed coding of the analyzed image. The last layer is the
reduced size characterization of the input image, representing the discriminant
features, that are used as the input attributes to the final recognizing system.

The rest of the report is organized as follows. In Section 2, I will give a
brief overview on PCA and Autoencoders with a focus on its application and
interpretations for facial images. The experimental results conducted on a
public database is given in Section 3. Finally, I draw a conclusion in Section
4.
